Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2019-01-07 reports/paper.tex
fix ordering of toms references
reformat

\documentclass[amsmath,amssymb,twocolumn,superscriptaddress,aps,prl]{revtex4-1}

@@

\author{A.~R.~McCluskey}
\thanks{A.R.M. and A.S.-F. contributed equally to this work.}
\email{a.r.mccluskey@bath.ac.uk}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}

\author{A.~Sanchez-Fernandez}
\thanks{A.R.M. and A.S.-F. contributed equally to this work.}
\altaffiliation[Present address ]{Department of Food Technology, Lund
University, SE-211 00 Lund, Sweden.}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}

@@

\author{T.~Arnold}
\email{tom.arnold@esss.se}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}
\affiliation{ISIS Neutron and Muon Source, Science and Technology Facilities
Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot, OX11 0QX,
UK}

@@

%\tableofcontents
Deep eutectic solvents (DES) are green, sustainable liquids that are obtained
through the combination of ionic species with compounds that act as hydrogen
bond donors, such as sugars, alcohols, amines, and carboxylic
acids\cite{Smith2014,Dai2013}.
The resulting extensive hydrogen bonding network is able to stabilise the
ionic species and allows the eutectic mixture to remain liquid at room
temperature\cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to
tune the solvent's physicochemical properties, such as
polarity\cite{Pandey2014}, viscosity and surface tension\cite{Smith2014},
network charge\cite{Zahn2016}, and
hydrophobicity\cite{Ribeiro2015,vanOsch2015}.
Recently DES have also been shown to exhibit a ``solvophobic'' effect through
the promotion of surfactant micelle
formation\cite{Sanchez-Fernandez2016,Arnold2015,Hsieh2018,Banjare2018},
phospholipid bilayer formation\cite{Bryant2017,Bryant2016,Gutierrez2009}, and
the ability to stabilise non-ionic polymer\cite{Sapir2016} and protein
conformations\cite{Sanchez-Fernandez2017}.

Phospholipid monolayers at the air/water interface have been widely studied
as simplistic models for biological membranes.
As such, they have been used to gain insight into many biological processes
that are technologically and medically relevant.
For example, investigations at the air/salt-water interface have identified
the importance that interactions between charged phospholipid heads and ions
present in solution have on the structure, monomer packing and stability of
the monolayer\cite{Mohwald1990,Kewalramani2010}.
However, the native environment for lipids in-vivo is far from a simple
aqueous solution.
In fact, it has been suggested\cite{Dai2013,Hammond2017} that DES might form
within the crowded cellular environment and could assist in solubilizing
biological species in an intermediate environment between that of the
hydrophobic phospholipid tails and highly polar water rich regions, thereby
assisting survival under extreme conditions such as freezing temperatures or
drought where the water content of cells is restricted.

@@

Recent developments in computational resources and software have enabled
powerful methodologies and algorithms to be harnessed by those from
non-expert backgrounds.
This has benefitted significantly from open-source software projects such as
the Python language\cite{vanRossum1995} and the Jupyter notebooks
framework\cite{Kluyver2016}.
In the area of NR and XRR, the landscape of data-analysis software is diverse,
with a range of software packages available from a variety of sources;
refnx\cite{Nelson2018}, MOTOFIT\cite{Nelson2006}, Rascal\cite{HughesRascal}
Aurore\cite{Gerelli2016}, Refl1D\cite{Kienzle2011}, and GenX\cite{Bjorck2007}.

The use of a Python library, such as refnx, enables the implementation custom
models that contain chemically-relevant information as well as the
application of probability distribution function (PDF) sampling techniques.
The Python library emcee\cite{Foreman-Mackey2013} offers refnx to access the
Goodman \& Weare Affine Invariant Markov chain Monte Carlo (MCMC) Ensemble
method\cite{Goodman2010}.
This allows the sampling of the high-dimensionality parameter space, relevant
in reflectomety analysis, in a Bayesian fashion, where the new samples are
generated with consideration of those sampled previously\cite{Sivia2006}.
Bayesian inference gives an understanding of the PDF for the fitted
parameters and therefore estimations of their inverse uncertainties and
inter-parameter correlations.

%
\begin{figure}[b]
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC) \cite{SI2018}.}
\end{figure}
%
We report on measurements of four phospholipids monolayers, namely
1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails),
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),
1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and
1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol)
(DMPG, C$_{14}$ tails), at the air-DES (1:2 choline choride:glycerol)
interface.
In contrast to many previous studies\cite{Mohwald1990,Kewalramani2010,
Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, we have developed
a chemically-consistent model that allows for the
co-refinement of reflectometry measurements at different surface pressure
and makes no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$ \cite{SI2018}.
Instead these parameters were allowed to vary for each lipid while being
constrained to be self-consistent over different surface pressures in the
same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for
DMPC, DMPG and DLPC.
This model was required because we cannot assume that the electrostatic
interactions between lipid head and the solvent are the same in the DES and
in water. This may therefore influence the effective head volume which means
we cannot rely on the literature values derived from measurements
on or in water \cite{SI2018}.
Furthermore, it is known that, on water, increased surface pressure and
associated LE-LC phase transitions lead to a compression of the lipid tail
volume\cite{Marsh2010,Small1984} and this compaction has not necessarily
been accounted for in the literature\cite{Campbell2018}.
Our approach avoids this issue by making no assumption about the molecular
volumes and only considering surface pressures that we believe to be in the
same phase.

%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:lipids}
\end{figure}
%
Our model has more variables than we can uniquely fit with limited data.
In reflectometry the usual approach to this problem is to collect more
equivalent dataset with varying deuteration (NR). For reasons of cost and
practicality this was not possible in this work.
The other approach is to constrain the model so that the number of fitted
variables are reduced, and is the approach taken here.
To do this we add the constraint that a single tail volume can be used for
all surface pressures. To justify this we need to be sure that the lipids
remain in the same phase. On water this is can be demonstrated with a
Langmuir isotherm.
However, while we have confidence that the individual surface pressures
measured were reliable, we were unable to collect consistent Langmuir
isotherm measurements, due to the high viscosity of the DES.
This inconsistency means that we cannot be confident in the phase of the
lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the
phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
DPPC was found to be in the LC phase and DMPC in the LE phase at room
temperature for the surface pressures measured (see Section \ref{sec:gixd}) \cite{SI2018}.
We assume that DMPG and DLPC are also in the LE phase since there is no
reason to believe that the phase behaviour in these systems differs
significantly from DMPC at the same temperature.

The model, based on the standard two-layer model widely used for lipids on
water, was first fitted to the experimental XRR data.
The associated SLD profiles are shown in Figure \ref{fig:lipids} while
Table \ref{tab:liptab} presents the results of the PDF for each of the
varying parameters; the tail tilt angle, $\theta_t$, the interfacial
roughness, $\sigma$, the head and tail volumes, $V_h$ and $V_t$ respectively,
and the head layer thickness, $d_h$.
In order to constrain the fitting process only $\theta_t$ and $\sigma$ were
allowed to vary independently of surface pressure. The other parameters,
$V_h$, $V_t$ and $d_h$ were fitted to a single value for all of the surface
pressures measured for each lipid.

%
\begin{table}
	\caption{\label{tab:liptab} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the XRR models, at the
  \SI{30}{\milli\newton\per\meter}. The values of $d_t$ were found from the
  appropriate values of $\theta_t$ using Eqn. \ref{equ:tl} and the values for
  $\phi_h$ were obtained from the appropriate use of Eqn. \ref{equ:phih} \cite{SI2018}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & DLPC & DMPC & DPPC & DMPG \\
    \hline
    $\theta_t$/\si{\degree} & \input{../output/dlpc/angle30.txt} &
    \input{../output/dmpc/angle30.txt} & \input{../output/dppc/angle30.txt} &
    \input{../output/dmpg/angle30.txt} \\
    $\sigma$/\si{\angstrom} & \input{../output/dlpc/rough30.txt} &
    \input{../output/dmpc/rough30.txt} & \input{../output/dppc/rough30.txt} &
    \input{../output/dmpg/rough30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} &
    \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} &
    \input{../output/dmpg/vt.txt} \\
    $V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} &
    \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} &
    \input{../output/dmpg/vh.txt} \\
    $d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} &
    \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} &
    \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh30.txt} &
    \input{../output/dmpc/solh30.txt} & \input{../output/dppc/solh30.txt} &
    \input{../output/dmpg/solh30.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dlpc/tail30.txt} &
    \input{../output/dmpc/tail30.txt} & \input{../output/dppc/tail30.txt} &
    \input{../output/dmpg/tail30.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%

For each lipid, the tail layer thickness, $d_t$, increases with chain length
and in general agrees with the literature for monolayers on
water\cite{Mohwald1990,Vaknin1991}:
For DMPC, \input{../output/dmpc/tail30.txt}\si{\angstrom} at
\SI{30}{\milli\newton\per\meter} in DES compared with
$d_t=\SI{15.8}{\angstrom}$ at
\SI{30}{\milli\newton\per\meter}\cite{Johnson1991} in water, and for
DPPC \input{../output/dppc/tail30.txt}\si{\angstrom} at
\SI{30}{\milli\newton\per\meter} in DES compared with
$d_t=\SI{16.7}{\angstrom}$ at
\SI{40}{\milli\newton\per\meter}\cite{Helm1987} in water.
In all cases the surface roughness was found to be relatively high compared
to water.
This is expected because the absolute surface tension of the DES is less
than water, while a previous XRR measurement of pure choline
chloride:glycerol suggests a roughness of
3.3\si{\angstrom}\cite{Sanchez-Fernandez2016}, significantly higher than
the capillary wave roughness of water of about 2.8\si{\angstrom}.

Figure \ref{fig:lipresults} shows the tail layer thickness variation with
surface pressure. For DPPC and DMPC a plateau in thickness is reached at
\SI{25}{\milli\newton\per\meter} and \SI{30}{\milli\newton\per\meter}
respectively. Presumably a similar plateau would be seen for DMPG and DLPC
at higher pressures.
This phenomenon has been noted before for DMPC\cite{Bayerl1990} and
DPPC\cite{Campbell2018} at the air-water interface.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{\small The PDFs of the head volume (left) and variation of
  $d_t$ (squares) and $\phi_h$ (circles) with surface pressure for each of
  the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The values of
  $d_t$ were found from the appropriate values of $\theta_t$ using
  Eqn. \ref{equ:tl} \cite{SI2018}.}
	\label{fig:lipresults}
\end{figure}
%

Figure \ref{fig:lipresults} also shows that for all four lipids there is a
decrease in head solvation, $\phi_h$, with surface pressure.
This is because an increase in the surface pressure means a reduction in the
free volume available between the lipid heads, which in turn forces solvent
molecules out of the layer, an effect that has also been observed on
water\cite{Bayerl1990}.

Our fits suggest lower lipid tail volumes than previous measurements using
other techniques (compare Tables \ref{tab:water} and \ref{tab:liptab}) \cite{SI2018}.
It is unlikely that this is a result of the DES subphase since the tails do
not directly interact with the solvent.
However this may be related to the compaction of the monolayer at elevated
surface pressures.
The optimal value of the tail volume for DPPC in the LC
phase\cite{Campbell2018} was found to be \SI{772}{\cubic\angstrom} at
\SI{35}{\milli\newton\per\meter}, which agrees well with the value of
\input{../output/dppc/vt.txt}\si{\cubic\angstrom} found in this work.
Similarly the tail volume for DMPS in the LE phase\cite{Campbell2018},
\SI{714}{\cubic\angstrom} at \SI{10}{\milli\newton\per\meter}, broadly
agrees with the values for
DMPC (\input{../output/dmpc/vt.txt}\si{\cubic\angstrom}) and DMPG
(\input{../output/dmpg/vt.txt}\si{\cubic\angstrom}) in this work.
The literature value\cite{Pan2012} quoted for DMPG (at a lower temperature) is
also similar to our result (but slightly smaller).
We find that the reduction is \SIrange{8}{12}{\percent} for DPPC, DMPC and
DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, in
good agreement with the maximum compression percentage of \SI{15}{\percent}
noted by Small \emph{et al.}\cite{Small1984}.
In general our results are at least self-consistent while comparisons to the
literature are favourable.

Figure \ref{fig:lipresults} shows the PDFs for the head volume for each of
the four lipids.
The three lipids with the PC head are consistent with values of around
\SI{330}{\cubic\angstrom}, regardless of hydrocarbon tail.
This agrees well with the values found for the same head in water
(Table \ref{tab:water}) \cite{SI2018}.
Interestingly, the volume for the PG head is similar to that for the PC head
with a value of \input{../output/dmpg/vh.txt}\si{\cubic\angstrom}, which is
significantly larger than values quoted in the literature for
DMPG\cite{Pan2012} (\SI{291}{\cubic\angstrom}) or POPG\cite{Kucerka2012}
(\SI{289}{\cubic\angstrom}).
This suggests that there may be some effect arising from the solvation in DES.

The major difference between the two heads is the fact PG head is negatively
charged whereas the PC head is zwitterionic (Figure \ref{fig:heads}).
It has been shown previously that the PC head is folded in
water\cite{Gilliams2016}, while the presence of strong electrostatic
interactions in DES is known to effect the structure of surfactants
micelles\cite{Sanchez-Fernandez2018}.
If we infer a similar folded structure for the PG head driven by a weaker
interaction between the alcohol and phosphate groups, then we may explain
the observed increase in volume to be due to unfolding of the PG head
because we expect that the DES provides a greater charge screening effect
than water.
If so then we would also expect an increase in the thickness of the head
layer in DES. This does seem to be the case: The PG head layer thickness was
found to be \SI[separate-uncertainty=true]{10.3\pm0.4}{\angstrom} at
\SI{22}{\milli\newton\per\meter} \cite{Clifton2012} and
\SI[separate-uncertainty=true]{9.7\pm1.0}{\angstrom} at
\SI{15}{\milli\newton\per\meter} \cite{Ciumac2017} from NR measurements at
the air-water interface.
For the PC head the stronger folding interaction, due to the formal charge
on the ammonium group, means that no unfolding effect is observed.

>>>>

\documentclass[amsmath,amssymb,twocolumn,superscriptaddress]{revtex4-1}

@@

\author{A.~R.~McCluskey}
%\thanks{A.R.M. and A.S.-F. contributed equally to this work.}
\email{a.r.mccluskey@bath.ac.uk/andrew.mccluskey@diamond.ac.uk}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}

\author{A.~Sanchez-Fernandez}
%\thanks{A.R.M. and A.S.-F. contributed equally to this work.}
\altaffiliation[Present address: ]{Department of Food Technology, Lund
University, SE-211 00 Lund, Sweden.}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}

@@

\author{T.~Arnold}
\email{tom.arnold@esss.se}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}
\affiliation{ISIS Neutron and Muon Source, Science and Technology Facilities
Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot, OX11 0QX,
UK}

@@

%\tableofcontents
\section{Introduction}
%
Deep eutectic solvents (DES) are green, sustainable liquids that are obtained through the combination of ionic species with compounds that act as hydrogen bond donors, such as sugars, alcohols, amines, and carboxylic acids \cite{Smith2014,Dai2013}.
The resulting extensive hydrogen bonding network is able to stabilise the ionic species and allows the eutectic mixture to remain liquid at room temperature \cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to tune the solvent's physicochemical properties, such as polarity \cite{Pandey2014}, viscosity and surface tension \cite{Smith2014}, network charge \cite{Zahn2016}, and hydrophobicity \cite{Ribeiro2015,vanOsch2015}.
Recently DES have also been shown to exhibit a ``solvophobic'' effect through the promotion of surfactant micelle formation \cite{Sanchez-Fernandez2016,Arnold2015,Hsieh2018,Banjare2018}, phospholipid bilayer formation \cite{Bryant2017,Bryant2016,Gutierrez2009}, and the ability to stabilise non-ionic polymer \cite{Sapir2016} and protein conformations \cite{Sanchez-Fernandez2017}.

Phospholipid monolayers at the air/water interface have been widely studied as simplistic models for biological membranes.
As such, they have been used to gain insight into many biological processes that are technologically and medically relevant.
For example, investigations at the air/salt-water interface have identified the importance that interactions between charged phospholipid heads and ions present in solution have on the structure, monomer packing and stability of the monolayer \cite{Mohwald1990,Kewalramani2010}.
However, the native environment for lipids in-vivo is far from a simple aqueous solution.
In fact, it has been suggested \cite{Dai2013,Hammond2017} that DES might form within the crowded cellular environment and could assist in solubilizing biological species in an intermediate environment between that of the hydrophobic phospholipid tails and highly polar water rich regions, thereby assisting survival under extreme conditions such as freezing temperatures or drought where the water content of cells is restricted.

@@

Recent developments in computational resources and software have enabled powerful methodologies and algorithms to be harnessed by those from non-expert backgrounds.
This has benefitted significantly from open-source software projects such as the Python language \cite{vanRossum1995} and the Jupyter notebooks framework \cite{Kluyver2016}.
In the area of NR and XRR, the landscape of data-analysis software is diverse, with a range of software packages available from a variety of sources; refnx \cite{Nelson2018,Nelson2018a}, MOTOFIT \cite{Nelson2006}, Rascal \cite{HughesRascal} Aurore \cite{Gerelli2016}, Refl1D \cite{Kienzle2011}, and GenX \cite{Bjorck2007}.

The use of a Python library, such as refnx, enables the implementation custom models that contain chemically-relevant information as well as the application of probability distribution function (PDF) sampling techniques.
The Python library emcee \cite{Foreman-Mackey2013} offers refnx to access the Goodman \& Weare Affine Invariant Markov chain Monte Carlo (MCMC) Ensemble method \cite{Goodman2010}.
This allows the sampling of the high-dimensionality parameter space, relevant in reflectomety analysis, in a Bayesian fashion, where the new samples are generated with consideration of those sampled previously \cite{Sivia2006}.
Bayesian inference gives an understanding of the PDF for the fitted parameters and therefore estimations of their inverse uncertainties and inter-parameter correlations.

\section{Experimental Section}
%
\subsection{Materials}
%
Choline chloride (\SI{99}{\percent} \%, Sigma-Aldrich) and glycerol (\SI{99}{\percent}, Sigma-Aldrich) d$_9$-choline chloride (\SI{99}{\percent}, \SI{98}{\percent} D, CK Isotopes) and d$_8$-glycerol (\SI{99}{\percent}, \SI{98}{\percent} D, CK Isotopes) were purchased and used without further purification.
The DES was prepared by mixing the precursors at a 1:2 ratio of chloine chlroide:glycerol, and heating at \SI{80}{\celsius} until a homogeneous, transparent liquid formed \cite{Smith2014}.
The solvent was equilibrated overnight at \SI{40}{\celsius} and subsequently stored under a dry atmosphere.
Due to the limited availability of the deuterated precursors, a fully protonated subphase (hDES) and a partially deuterated subphase (hdDES) were prepared and used during the neutron reflectometry (NR) experiment.
The partially deuterated subphase was prepared using the following mixtures of precursors: \SI{1}{\mole} of \SI{0.38}{\mole} fraction of h-choline chloride/\SI{0.62}{\mole} fraction of d-choline chloride; and \SI{2}{\mole} of \SI{0.56}{\mole} fraction of h-glycerol/\SI{0.44}{\mole} fraction of d-glycerol.
The solvent was subsequently prepared following the procedure discussed above.

The water content of the DES was determined before and after each experiment by Karl-Fischer titration (Mettler Toledo DL32 Karl-Fischer Coulometer, Aqualine Electrolyte A, Aqualine Catholyte CG A) in order to ensure water presence was kept to a minimum.
Those measurements showed that the water content of the solvent was kept below \SI{0.3}{wt/\percent} during all the experimental procedures presented here, which we assume to be negligible and have to little impact on the characteristics of the DES \cite{Hammond2016,Hammond2017}.

1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails, \SI{>99}{\percent}), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails, \SI{>99}{\percent}), and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG, C$_{14}$ tails, \SI{>99}{\percent}) were supplied by Avanti Polar Lipids and, 2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails, \SI{>99}{\percent}) was supplied by Sigma Aldrich and all were used without further purification.
Deuterated versions of DPPC (d$_{62}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) and DMPC (d$_{54}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification.
These phopholipids were dissolved in chloroform (\SI{0.5}{\milli\gram\per\milli\liter}) at room temperature.
PC indicates the molcule contains a phosphocholine head component, where PG contains a phosphatidylglycerol head component, these are shown in Figure \ref{fig:heads}.
%
\begin{figure}
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC) \cite{SI2018}.}
\end{figure}
%

In the XRR experiment, sample preparation was performed in situ using the standard method for the spreading of insoluble monolayers on water: a certain amount of the phospholipid solution was spread onto the liquid surface in order to provide a given surface concentration.
After the evaporation of the chloroform, it is assumed that the resulting system is a solvent subphase with a monolayer of phospholipid at the interface.
Surface concentration was modified by closing and opening the PTFE barriers of a Langmuir trough.
In order to minimise the volumes used in the NR experiment (to keep the cost of deuterated compounds to a manageable level) it was not possible to use a Langmuir trough.
Instead, small Delrin adsorption troughs were used that did not have controllable barriers.
So, although the surface coverage was nominally the same as used in the X-ray studies, the lack of precise control over the surface pressure meant that it was not appropriate to co-refine XRR and NR contrasts together.

\subsection{Methods}
%
XRR measurements were taken on I07 at Diamond Light Source, at \SI{12.5}{\kilo\electronvolt} photon energy using the double-crystal deflector \cite{Arnold2012}.
The reflected intensity was measured in a momentum transfer range from \SIrange{0.018}{0.7}{\per\angstrom}.
The data were normalised with respect to the incident beam and the background was measured from off-specular reflection and subsequently subtracted.
Samples were equilibrated for at least one hour and preserved under an argon atmosphere to minimise the adsorption of water by the subphase.
XRR data were collected for each of the lipids, DLPC, DMPC, DPPC and DMPG at four surface pressures (DLPC: \SIlist[list-units = single]{20;25;30;35}{\milli\newton\per\meter}, DMPC: \SIlist[list-units = single]{20;25;30;40}{\milli\newton\per\meter}, DPPC: \SIlist[list-units = single]{15;20;25;30}{\milli\newton\per\meter}, DMPG: \SIlist[list-units = single]{15;20;25;30}{\milli\newton\per\meter}, as measured with an aluminium Wilhelmy plate; all measurements were conducted at \SI{22}{\celsius}.
The aluminium Wilhelmy plate was used over a traditional paper plate due to the low wetability of paper by the DES.

The NR experiments were performed on FIGARO at the Institut Laue-Langevin using the time-of-flight method \cite{Campbell2011}.
Data at two incident angles of \SI{0.62}{\degree} and \SI{3.8}{\degree} were measured to provide a momentum transfer range from \SIrange{0.005}{0.18}{\per\angstrom}.
Two surface pressures for each system and contrast was measured (DMPC: \SIlist[list-units = single]{20;25}{\milli\newton\per\meter}, DPPC: \SIlist[list-units = single]{15;20}{\milli\newton\per\meter}).
Similar to the X-ray procedure, samples were given enough time to equilibrate (at least two hours), kept under an inert atmosphere, and all measurements were conducted at \SI{22}{\celsius}.

\subsection{Data analysis}
%
The use of XRR and NR to analyse the structure of phospholipids on the surface of water is well documented \cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}.
The models used in the rationalisation of XRR and NR data have varied significantly in numbers of layers present, use of interfactial roughness, and the parameterisation of the physical constraints applied.
Frequently, these physical constraints include the volumes of the phospholipid head and tail components, using values taken from other techniques, such as those shown in Table \ref{tab:water}.
Additionally, a recent evaluation of the applicability of different models for surfactant and phospholipid monolayers from the NR perspective has been published \cite{Campbell2018}, that suggests possible oversights in the modelling of NR data.
%
\begin{table*}
	\caption{\label{tab:water} Lipid component volumes extracted from different literature sources.
  $V_l$ corresponds to the total lipid volume, MD to molecular dynamics simulation, WAXS to wide-angle X-ray scattering, NB to neutral buoyancy and DVTD to differential vibrating tube densimetry. $^a$
  The values for the head component in Kucerka \emph{et al.} \cite{Kucerka2004}, were taken from Balgav\'{y} \emph{et al} \cite{Balgavy2001}.}
  \begin{ruledtabular}
	\begin{tabular*}{\textwidth}{l|lll|ll|ll|l|l}
    Lipid & DPPC & & & DMPC & & DLPC & & DMPG & POPG \\
    \hline
    Reference & \cite{Armen1998} & \cite{Sun1994} &
    \cite{Kucerka2004,Balgavy2001}$^a$ & \cite{Armen1998}] & \cite{Kucerka2004,Balgavy2001}$^a$ & \cite{Armen1998} & \cite{Kucerka2004,Balgavy2001}$^a$ & \cite{Pan2012} &
    \cite{Kucerka2012} \\
    \hline
    $V_l$/\AA$^3$ & $1287.3\pm25.5$ & $1148\pm2$ & $1264.2\pm32.1$ &
    $1172.5\pm25.1£$ & $1155.4\pm30.0£$ & $1057.7\pm24.7$ & $1046.6\pm28.0$ &
    $1011.4$ & $1203$ \\
    $V_t$/\AA$^3$ & $966.4\pm5.4$ & $829\pm4$ & $924.7\pm17.6$ &
    $851.5\pm5.0£$ & $815.9\pm15.5£$ & $736.8\pm4.6$ & $707.1\pm13.5$ &
    $720.4$ & $914$ \\
    $V_h$/\AA$^3$ & $320.9\pm20.1$ & $319\pm6$ & $339.5\pm14.5$ &
    $320.9\pm20.1$ & $339.5\pm14.5$ & $320.9\pm20.1$ & $339.5\pm14.5$ &
    $291.0$ & $289$ \\
    Method & MD & WAXS & NB & MD & NB & MD & NB & DVTD & MD \\
    T/$^\circ$C & 50 & 24 & 30 & 50 & 30 & 50 & 30 & 20 & 25 \\
	\end{tabular*}
  \end{ruledtabular}
\end{table*}
%

In Table \ref{tab:water}, there appears to be a general consensus that the component volume for the phosphocholine (PC) head is around \SIrange{320}{360}{\angstrom\cubed}, while the phosphatidylglycerol (PG) head is in the range \SIrange{289}{291}{\angstrom\cubed}.
However, it is not clear is the head component volumes from the literature, that are derived from water-based measurements, will be appropriate for this work, which involves a non-aqueous solvent.
The charged nature of the zwitterionic or anionic lipid heads may have different interactions with the polar, but neutral water as compared to the charged DES components \cite{Sanchez-Fernandez2018}, which will affect the phospholipid head component volume.
Therefore, herein we apply a chemically-consistent model that allows for the co-refinement of reflectometry measurements at different surface pressure, and makes no assumption of the component volume for the lipid head, $V_h$, or tail, $V_t$ \cite{SI2018}.
Instead, these parameters were allowed to vary for each lipid while being constrained to be self-consistent over different surface pressures in the same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for DMPC, DMPG, and DLPC.
Furthermore, it is known that, on water, increase surface pressure and the associated LE-LC phase transitions lead to a compression of the lipid tail volume \cite{Marsh2010,Small1984}, and this compaction has not necessarily been accounted for in the literature \cite{Campbell2018}.
This model avoids this issue by making no assumption about the molecular volumes and only considers surface pressures that we believe to be in the same phase.

Our chemically-consistent model has been implemented in the Python library refnx \cite{Nelson2018,Nelson2018a}.
This software allows for the inclusion of a custom model to be defined, from which parameters feed into the Abel\`{e}s refelectivity model (a model that is widely used to calculate reflectivity \cite{Abeles1950,Parratt1954}).
This custom model, along with a series of Jupyter notebooks showing, in full, the analysis performed, can be found in the ESI and is available under a CC BY-SA 4.0 licence \cite{SI2018}.
